A Longitudinal Analysis of the Privacy Paradox

### 2020-05-07 manuscript.RmD
- language updates - shortened manuscript

  - id            : "1"
    institution   : "University of Hohenheim"
  - id            : "2"
    institution   : "Johannes Gutenberg University Mainz"
author_note: |
  All authors contributed extensively to the work presented in this paper. TD, PM, & ST designed the study; PM supervised the data collection; PM administered the data importation; TD & PM wrote the code, ran the models, and analyzed the output data; TD wrote major parts of manuscript, and PM & ST contributed individual sections and comments; ST supervised the project and wrote the grant application (in 2012). 
  The authors declare no competing interests.
  This research was funded by the German Federal Ministry of Education and Research (BMBF) Grant 16KIS0094, awarded to Sabine Trepte.

  This manuscript features a companion website that includes detailed summaries of the statistical results, the code, additional analyses, and a reproducible version of the manuscript (https://tdienlin.github.io/privacy-paradox-longitudinal). The data can be downloaded from http://dx.doi.org/10.7802/1937. 
abstract: |
  The privacy paradox states that people's concerns about online privacy are unrelated to their online sharing of personal information. 
  Using a representative sample of the German population, which includes `r knitr::load_cache(label = "load-data", object = "n_complete")` respondents who were interviewed at three waves separated by 6 months, we investigate the privacy paradox from a longitudinal perspective, differentiating between-person relations from within-person effects.
  Results of a cross-lagged panel model with random intercepts revealed that people who were more concerned about their online privacy than others also shared slightly less personal information online and had substantially more negative attitudes toward information sharing (between-person level). 
  Next, people who were more concerned than usual also shared slightly less information than usual (within-person level). 
  At the same time, we found no long-term effects of privacy concerns on information sharing or attitudes 6 months later.
  Together, the results provide further evidence against the privacy paradox.
keywords          : "privacy paradox, privacy concerns, information sharing, longitudinal analysis, structural equation modeling"
wordcount         : "`r wordcountaddin::word_count('manuscript.Rmd')`"
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
bibliography      : 
  - [bibliography/references.bib]
  - [bibliography/r-references.bib]
lang              : "english"
class             : "man"
mask              : no
# to compile as pdf
# (a) move file "_site.yml" to folder "tmp"
# (b) uncomment the following lines; comment lines after "output: html_document ..."
# output            : papaja::apa6_pdf
# header-includes:
#   - \setlength{\parskip}{0em}
#   - \raggedbottom
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    toc_depth: 2
---

@@

# load r-bib
r_citations <- cite_r("bibliography/r-references.bib", 
                      footnote = FALSE, 
                      pkgs = c("GGally", "ggplot2", "lavaan", "lme4", "magrittr", 
                               "MissMech", "MVN", "papaja", "psych", "pwr",  
                               "semTools", "sjstats", "tidyverse", "td"), 
                      withhold = FALSE)
```

@@

The privacy paradox states that the information disclosure practices of Internet users are problematic: 
Although many people are concerned about their online privacy, they still tend to share plenty of personal information on the web [e.g., @acquistiLossesGainsHyperbolic2003]. 
The privacy paradox and its underlying theoretical conundrum is hence of considerable interest to society---it is discussed in newspapers [@freanFindingSolutionsPrivacy2017], Wikipedia [@wikipediaPrivacyConcernsSocial2018], designated websites [@newyorkpublicradioPrivacyParadox2018], books [@treptePrivacyOnlinePerspectives2011], and top-tier academic journals [@acquistiPrivacyHumanBehavior2015].
If the privacy paradox really exists, it should inspire worry: 
It would suggest that online behavior is irrational and that people are revealing too much of their personal lives to unknown third parties, which can foster potentially unintended consequences such as commodification [@sevignaniPrivacyCapitalismAge2016], inappropriate recontextualization [@boydTakenOutContext2008], or increased predictability of future behavior [@bagrowInformationFlowReveals2019].
In conclusion, understanding why people disclose information online and whether this is paradoxical or not represents an important challenge for scholars in the social sciences.

However, current research on the privacy paradox has one major limitation: 
To the best of our knowledge, all empirical studies conducted so far have investigated the privacy paradox from a between-person (i.e., interpersonal) perspective.
By employing empirical tests relations between people (e.g., cross-sectional questionnaires analyzed with multiple regression or Pearson correlations), studies have analyzed whether people who are more concerned than _others_ also share less personal information than _others_.
Although such a between-person perspective is interesting and represents a viable first step, it cannot make informed claims regarding causality.
The privacy paradox, however, actually implies a within-person (i.e., intrapersonal) causal perspective:
Does a person, if _he or she_ becomes more concerned about online privacy, then also share less personal information?
This mismatch is problematic because although between-person relations are, except for some special cases, a _necessary_ condition for causal within-person effects, they are by no means a _sufficient_ one.
For example, it could be that the between-person relation is determined by another third variable. 
Hence, as the next step in investigating the privacy paradox and to better understand the intrapersonal causal relation between privacy concerns and information sharing, it is necessary to conduct studies with within-person designs.

With this study we therefore aim to answer four major questions. 
First, on a between-person level, how are concerns about online privacy related to the online sharing of personal information? 
Second, on a within-person level, does information sharing decrease when concerns increase?
Third, what are the potential long-term effects, are changes in concerns related to changes in information sharing 6 months later and/or vice versa? 
Fourth, what is the role of privacy attitudes, do they mediate the relation between privacy concerns and information sharing?
To best answer and contextualize these questions, we first provide an in-depth theoretical analysis of the privacy paradox, after which we present the empirical results of a longitudinal panel study.

@@

@acquistiLossesGainsHyperbolic2003 were among the first to argue that the online disclosure of personal information is paradoxical. 
"Experiments reveal that very few individuals actually take any action to protect their personal information, even when doing so involves limited costs" (p.1). 
Three years later, @barnesPrivacyParadoxSocial2006 discussed the behavior of young people using SNSs, popularizing the term _privacy paradox_. 
Barnes listed six aspects of online behavior that she considered to be particularly paradoxical: (a) illusion of privacy, (b) high quantity of information sharing, (c) attitude behavior discrepancy, (d) lack of privacy concerns, (e) lack of privacy literacy, and (f) fabrication of false information.
@norbergPrivacyParadoxPersonal2007 was one of the first to explicitly analyze the privacy paradox empirically.
The study finds a mismatch between concerns and behavior, which is aligned with several other experimental studies conducted at the time [beresfordUnwillingnessPayPrivacy2012; hannOvercomingOnlineInformation2007; hubermanValuatingPrivacy2005].

While there are various understandings and operationalizations of the privacy paradox [@kokolakisPrivacyAttitudesPrivacy2017], subsequent research on the privacy paradox arguably focused on Barnes's third tenet, the attitude-behavior discrepancy. 
Whereas some studies reported that privacy concerns were not significantly related to the disclosure of personal information [e.g., @grossInformationRevelationPrivacy2005; @taddickenPrivacyParadoxSocial2014; @tufekciCanYouSee2008], which lends credence to the privacy paradox, 
a different set of studies showed relations that were statistically significant [e.g., @dienlinPrivacyParadoxRelic2015; @heirmanPredictingAdolescentsDisclosure2013; @walraveConnectingProtectingComparing2012], which refutes the privacy paradox.

Notably, in a parallel line of research other studies have also analyzed the relation between privacy concerns and information sharing; 
however, the term privacy paradox has often not been mentioned explicitly. 
Instead, studies have referred to the so-called _privacy calculus_. 
The privacy calculus states that sharing personal information is affected by both the respective costs and the anticipated benefits [@culnanInformationPrivacyConcerns1999].
By now, several studies have found empirical support for the privacy calculus in various online contexts [e.g., @bolUnderstandingEffectsPersonalization2018; @dienlinExtendedPrivacyCalculus2016; @krasnovaOnlineSocialNetworks2010].

@baruhOnlinePrivacyConcerns2017 published the first empirical meta-analysis on the relations between privacy concerns and various forms of social media use (e.g., information sharing or SNS usage). 
On the basis of 37 studies, @baruhOnlinePrivacyConcerns2017 found a small and significant statistical relation between concerns about online privacy and online information sharing (_r_ = -.13, 95% CI [-.07, -.18]). 
Another more recent meta analysis by @yuMetaanalysisExplorePrivacy2020 also finds a significant bivariate relation between privacy concerns and information sharing, although smaller (_r_ = -.06, 95% CI [-.01, -.12]]). 
There also exist several systematic literature reviews on the privacy paradox [@barthPrivacyParadoxInvestigating2017; @gerberExplainingPrivacyParadox2018; @kokolakisPrivacyAttitudesPrivacy2017]
@kokolakisPrivacyAttitudesPrivacy2017 conclude that "the dichotomy between privacy attitude and behaviour should not be considered a paradox anymore." (p. 130)
However, the authors also note that the privacy paradox is a "complex phenomenon that has not been fully explained yet".
@barthPrivacyParadoxInvestigating2017 are more skeptical, and argue that "attempts to theoretically explain and practically solve the problem of the privacy paradox are still scarce and we feel the subject deserves far more research attention" (p. 1052).

In conclusion, large parts of the current literature suggests that a significant relation between concerns about online privacy and the online sharing of personal information exists and that it is small---which speaks against the privacy paradox.

@@

Privacy is defined as the "[...] voluntary and temporary withdrawal of a person from the general society through physical or psychological means [...]" [@westinPrivacyFreedom1967, 7]. 
Hence, privacy captures aspects of both volitional _control_ and social _separateness_.
Several dimensions of privacy have been proposed:
For example, it is possible to distinguish a vertical and a horizontal level;
whereas the vertical level captures privacy from authorities, institutions, or companies, horizontal privacy addresses privacy from peers, colleagues, or other people [@masurSituationalPrivacySelfdisclosure2018].
When it comes to concerns specifically, interestingly they do not seem to be established as a stand-alone theoretical concept in psychology [@colmanDictionaryPsychology2015].
Lexically, concerns are defined as a "marked interest or regard usually arising through a personal tie or relationship" that also reflect an "uneasy state of blended interest, uncertainty, and apprehension" [@merriam-websterDefinitionConcern2018].
Concerns therefore represent both a latent _motivation_ (or increased attention) and a negatively valenced _emotion_ (or affective condition).
As such, concerns are not the result of a deliberate explicit cognition.
Instead, they stem from an automatic implicit perception.
Taken together, concerns about online privacy represent how much an individual is motivated to focus on his or her control over a voluntary withdrawal from other people or societal institutions on the Internet, accompanied by an uneasy feeling that his or her privacy might be threatened.
As a theoretical construct, privacy concerns can hence be categorized as an affective motivational disposition.
As such, there are many similarities with other concepts, which includes emotions (e.g., fear, anxiety), moods (e.g., dismay, fatigue), attitudes (risk perception, approval), values (e.g., autonomy, freedom), personality traits (e.g., introversion, risk avoidance), and even physiological activation (e.g., attention, arousal).

The online sharing of personal information, on the other hand, captures how much person-related information people share when they use the Internet, which includes, for example, information about their age, sex, name, address, health, or finances.
Information sharing can be differentiated from communication and self-disclosure:
Communication is broad, because it comprises all verbal and nonverbal information that is emitted [e.g., @watzlawickPragmaticsHumanCommunication2011].
Self-disclosure is narrow, because it focuses on deliberate revelations about the true self to others [e.g., @jourardTransparentSelf1964].
Information sharing is even more specific, because it addresses only person-related information but ignores other types of self-disclosure such as personal fears, values, or plans.

@@

Currently, there seems to be a lack of studies that explicitly analyze how behavior is affected by _concerns_.
Fortunately, however, we know much about the behavioral effects of related concepts such as attitudes or fears, which all can affect behavior profoundly [@fishbeinPredictingChangingBehavior2010; @rogersCognitivePhysiologicalProcesses1983].
Emotions, perhaps the concept most closely related to concerns, have a particularly strong effect on behavior. 
By causing fight, flight, or freeze reactions, they are a primordial trigger of behavior and an adaptive mechanism fostering the evolution of a species [@dolanEmotionCognitionBehavior2002].
With their direct link to the amygdala, emotions can trigger reactions already subcortically [i.e., without activation of the more recently evolved cortical structures; @dolanEmotionCognitionBehavior2002].

Also empirically, concerns have been shown to affect behavior: 
People more concerned about the environment show more environment-related behaviors [@bambergHowDoesEnvironmental2003]; people more concerned about their appearance consume fewer calories [@hayesConcernAppearanceHealth1987];
People more concerned about their bodies engage in more physical exercise [@reelRelationsBodyConcerns2007].
Taken together, it is reasonable to expect that also concerns about online privacy should reflect in the online sharing of personal information.

At the same time, there are some factors that are likely to diminish the relation.
Most prominently, there is the so-called _attitude behavior gap_ [@fishbeinPredictingChangingBehavior2010], which states that people sometimes act against their own attitudes. 
Not everyone concerned about his or her physical health exercises regularly.
The explanation is simple.
Other factors such as subjective norms and perceived behavioral control also determine behavior [@ajzenIntentionsActionsTheory1985], which automatically reduces the impact of attitudes or concerns. 
Specifically, two of the most influential factors that affect online information sharing are (a) strong subjective norms to participate online [@heirmanPredictingAdolescentsDisclosure2013] and (b) the manifold benefits that accrue from participation [@krasnovaOnlineSocialNetworks2010]. 
In other words, users often prioritize social support, special offers, or improved services.
@trepteRiskyBehaviorsHow2014 listed several factors that can additionally attenuate the relation:
lack of strength of concerns,
absence of negative personal experiences, 
and situational constraints due to social desirability.
In conclusion, also in the context of the privacy paradox we should not expect to find a perfect relation between attitudes and behaviors, but rather a small to moderate relation.

Finally, there are also some methodological reasons that explain why some studies did not detect statistically significant relations.
In general, researchers are confronted with the _Duhem-Quine problem_, which holds that it is impossible to test theories in isolation, because empirical tests always rely on auxiliary assumptions [@dienesUnderstandingPsychologyScience2008]. 
In other words, if a psychological experiment fails, we do not know whether the theory is wrong or the questionnaire subpar. 
This tenet is particularly relevant for the privacy paradox: 
Detecting statistical significance for small effects---and in this case, we should expect to find small effects---is more challenging because it means that large samples are necessary to guarantee sufficient statistical power.[^power]
Precisely, in order to be capable of detecting a correlation between privacy concerns and information sharing in 95% of all cases, which @baruhOnlinePrivacyConcerns2017 estimated to be _r_ = -.13, we need a sample of _N_ = `r pwr.r.test(r = -.13, sig.level = .05, power = .95)$n %>%  round(0)` people. 
The reality, however, looks different: 
In their meta-analysis, @baruhOnlinePrivacyConcerns2017 reported a median sample size of _N_ = 300, which can explain why several studies did not find significant effects.

@@

In conclusion, in line with prior research [@baruhOnlinePrivacyConcerns2017] and the within-person rationales that we have presented above, we expect to find a small significant relation between privacy concerns and information sharing, both on the between-person level and the within-person level.

@@

Although short-term effects are likely, it is unclear whether there will be any long-term effects.
First, when analyzing potential long-term effects, it is important to choose an interval that is both plausible and relevant.
For example, it makes a difference whether the effects of alcohol consumption on driving performance are tested 1 minute, 1 hour, or 1 day after consumption. 
One factor that determines an interval's optimal length is the variable's stability [@dormannOptimalTimeLags2015].
Privacy concerns and privacy attitudes are predominantly trait-like constructs with high stabilities, which is why they necessitate longer intervals.
Other studies with comparable research questions have used an interval of 6 months [e.g., @valkenburgEffectsInstantMessaging2009], which we consider to be plausible also in this case.

@@

At the same time, the adverse effect of information sharing on privacy concerns seems also plausible, with two potential outcomes. 
On the one hand, the long-term relation could be negative: 
If people start to share more information online, they might become increasingly aware that their privacy is at risk, which might stir concern.
On the other hand, the long-term relation might also be positive, because when people share more personal information online, they might become accustomed to doing so, which might reduce concern [for example, due to the mere exposure effect; @zajoncAttitudinalEffectsMere1968].
Finally, there could also be no long-term relation after all. 
For example, people might have already become used to sharing information online, which might stifle any further cognitive or emotional processing -- a rationale central to the observation of so-called _privacy cynicism_ [e.g., @hoffmannPrivacyCynicismNew2016].

@@

It has been argued that privacy attitudes could 'bridge the gap' between concerns and information sharing [e.g., @dienlinPrivacyParadoxRelic2015]. 
In contrast to privacy concerns, privacy attitudes capture a more explicit, fluctuating cognitive appraisal. 
Although both variables are related to information disclosure, attitudes are likely the better predictor. 
This reasoning follows the rational choice paradigm [@simonBehavioralModelRational1955], which maintains that behavior is always, at least in part, influenced by convictions, attitudes, and cost-benefit analyses. 
<!-- Likewise, also evolutionary psychology suggests that although emotions guide behavior in the long run, it is more adaptive if behavior instead reflects a cognitive appraisal;  -->
<!-- a tenet arguably reflected best by the phylogenetic development of the neocortex or, if you want, the general dominion of the human race. -->
Also empirically, a study of 1,042 youths from Belgium found that the relation between privacy attitudes and disclosure of personal information was strong (_r_ = .56), whereas the relation between privacy concerns and disclosure was only moderate [_r_ = -.29; @heirmanPredictingAdolescentsDisclosure2013].

@@

Concerning the potential long-term relations of privacy attitudes, we are confronted with the same situation mentioned above:
Given that we are not aware of research on long-term relations, several scenarios seem plausible.
For example, attitudes could either have long-term relations or not, and information sharing could either foster privacy attitudes or diminish them.

@@

This study is part of a large-scale project which investigates the development of privacy and self-disclosure, including several other variables. 
Other publications linked to the project can be accessed at https://osf.io/y35as/.
The data come from a longitudinal paper-and-pencil questionnaire study, in which a representative sample of the German population (16 years and older) was surveyed on overall five occasions.
The data can be downloaded from http://dx.doi.org/10.7802/1937.

>>>>

  - id            : "1"
    institution   : "University of Hohenheim"
  - id            : "2"
    institution   : "Johannes Gutenberg University Mainz"
author_note: |
  All authors contributed extensively to the work presented in this paper. TD, PM, & ST designed the study; PM supervised the data collection; PM administered the data importation; TD & PM wrote the code, ran the models, and analyzed the output data; TD wrote most parts of manuscript, and PM & ST contributed individual sections and comments; ST supervised the project and wrote the grant application (in 2012). 
  The authors declare no competing interests.
  This research was funded by the German Federal Ministry of Education and Research (BMBF) Grant 16KIS0094, awarded to Sabine Trepte.

  This manuscript features a companion website that includes detailed summaries of the statistical results, the code, additional analyses, and a reproducible version of the manuscript (https://tdienlin.github.io/privacy-paradox-longitudinal). The data can be downloaded from http://dx.doi.org/10.7802/1937. 
abstract: |
  The privacy paradox states that people's concerns about online privacy are unrelated to their online sharing of personal information. 
  Using a representative sample of the German population, which includes `r knitr::load_cache(label = "load-data", object = "n_complete")` respondents who were interviewed at three waves separated by 6 months, we investigate the privacy paradox from a longitudinal perspective. We differentiate between-person relations from within-person effects.
  Results of a cross-lagged panel model with random intercepts revealed that people who were more concerned about their online privacy than others also shared slightly less personal information online and had substantially more negative attitudes toward information sharing (between-person level). 
  People who were more concerned than usual also shared slightly less information than usual (within-person level). 
  We found no long-term effects of privacy concerns on information sharing or attitudes 6 months later.
  The results provide further evidence against the privacy paradox, but more research is needed to better understand the variables' potential causal relation.
keywords          : "privacy paradox, privacy concerns, information sharing, longitudinal analysis, structural equation modeling"
wordcount         : "`r wordcountaddin::word_count('manuscript.Rmd')`"
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
bibliography      : 
  - [bibliography/references.bib]
  - [bibliography/r-references.bib]
lang              : "english"
class             : "man"
mask              : no
# to compile as pdf
# (a) move file "_site.yml" to folder "tmp"
# (b) uncomment the following lines; comment lines after "output: html_document ..."
# output            : papaja::apa6_pdf
# header-includes:
#   - \setlength{\parskip}{0em}
#   - \raggedbottom
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    toc_depth: 2
---

@@

# load r-bib
r_citations <- cite_r("bibliography/r-references.bib", 
                      footnote = FALSE, 
                      pkgs = c("GGally", "ggplot2", "lavaan", 
                               "MissMech", "MVN", "psych", "pwr",  
                               "semTools", "sjstats"), 
                      withhold = FALSE)
```

@@

The privacy paradox states that the information disclosure of Internet users is problematic: 
Although many people are concerned about their online privacy, they still share plenty of personal information on the web [e.g., @acquistiLossesGainsHyperbolic2003]. 
The privacy paradox and its underlying theoretical conundrum is hence of considerable interest to society---it is discussed in newspapers [@freanFindingSolutionsPrivacy2017], Wikipedia entries  [@wikipediaPrivacyConcernsSocial2018], designated websites [@newyorkpublicradioPrivacyParadox2018], books [@treptePrivacyOnlinePerspectives2011], and top-tier academic journals [@acquistiPrivacyHumanBehavior2015].
If the privacy paradox really exists, it should inspire worry: 
It would suggest that online behavior is irrational and that people are revealing too much of their personal lives to unknown third parties, which can cause various problems [e.g., @sevignaniPrivacyCapitalismAge2016].
Understanding why people disclose information online and whether or not this is paradoxical therefore represents an important challenge.

However, current research on the privacy paradox has one major limitation: 
To the best of our knowledge, most empirical studies conducted so far have investigated the privacy paradox from a between-person (i.e., interpersonal) perspective.
By employing empirical tests of relations between people (e.g., cross-sectional questionnaires analyzed with multiple regression or Pearson correlations), studies have analyzed whether people who are more concerned than _others_ also share less personal information than _others_.
Although such a between-person perspective is interesting and represents a viable first step, it cannot make informed claims regarding causality.
The privacy paradox, however, implies a causal perspective:
Does a person, if _he or she_ becomes more concerned about online privacy, then also share less personal information?
This mismatch is problematic because although between-person relations are, except for some special cases, a _necessary_ condition for causal within-person effects, they are by no means a _sufficient_ one.
For example, it could be that the between-person relation is determined other third variables. 
Hence, as the next step in investigating the privacy paradox and to better understand the causal relation between privacy concerns and information sharing, it is necessary to conduct studies with within-person designs.

With this study we aim to answer four major questions. 
First, on a between-person level, how are concerns about online privacy related to the online sharing of personal information? 
Second, on a within-person level, does information sharing decrease when concerns increase?
Third, what are the potential long-term effects, are changes in concerns related to changes in information sharing 6 months later and/or vice versa? 
Fourth, what is the role of privacy attitudes, do they mediate the relation between privacy concerns and information sharing?
To best answer and contextualize these questions, we first provide an in-depth theoretical analysis of the privacy paradox, after which we present the empirical results of a longitudinal panel study.

@@

@acquistiLossesGainsHyperbolic2003 were among the first to argue that the online disclosure of personal information is paradoxical. 
"Experiments reveal that very few individuals actually take any action to protect their personal information, even when doing so involves limited costs" (p.1). 
Three years later, @barnesPrivacyParadoxSocial2006 discussed the behavior of young people on SNSs, popularizing the term _privacy paradox_. 
Barnes listed six aspects of online behavior that she considered to be particularly paradoxical: (a) illusion of privacy, (b) high quantity of information sharing, (c) attitude behavior discrepancy, (d) lack of privacy concerns, (e) lack of privacy literacy, and (f) fabrication of false information.
@norbergPrivacyParadoxPersonal2007 were one of the first to empirically analyze the privacy paradox explicitly.
The study found a mismatch between concerns and behavior, which is aligned with several other experimental studies conducted at the time [@beresfordUnwillingnessPayPrivacy2012; @hannOvercomingOnlineInformation2007; @hubermanValuatingPrivacy2005].

While there are various understandings and operationalizations of the privacy paradox [@kokolakisPrivacyAttitudesPrivacy2017], subsequent research focused on Barnes's third tenet, the attitude-behavior discrepancy. 
Whereas some studies reported that privacy concerns were not significantly related to the disclosure of personal information [e.g., @grossInformationRevelationPrivacy2005; @taddickenPrivacyParadoxSocial2014; @tufekciCanYouSee2008], which lends credence to the privacy paradox, 
a different set of studies showed significant relations [e.g., @dienlinPrivacyParadoxRelic2015; @heirmanPredictingAdolescentsDisclosure2013; @walraveConnectingProtectingComparing2012], which refutes the privacy paradox.

Notably, in a parallel line of research other studies have also analyzed the relation between privacy concerns and information sharing.
However, the term privacy paradox was often not mentioned explicitly. 
Instead, studies have referred to the so-called _privacy calculus_, 
which states that sharing personal information is affected by both the respective costs and the anticipated benefits [@culnanInformationPrivacyConcerns1999].
By now, several studies have found empirical support for the privacy calculus in various online contexts [e.g., @bolUnderstandingEffectsPersonalization2018; @dienlinExtendedPrivacyCalculus2016; @krasnovaOnlineSocialNetworks2010].

@baruhOnlinePrivacyConcerns2017 published the first empirical meta-analysis on the relations between privacy concerns and various forms of social media use (e.g., information sharing or SNS usage). 
On the basis of 37 studies, @baruhOnlinePrivacyConcerns2017 found a small and significant statistical relation between concerns about online privacy and online information sharing (_r_ = -.13, 95% CI [-.07, -.18]). 
Another more recent meta analysis by @yuMetaanalysisExplorePrivacy2020 also finds a significant bivariate relation between privacy concerns and information sharing, albeit smaller (_r_ = -.06, 95% CI [-.01, -.12]]). 
There also exist several systematic literature reviews on the privacy paradox [@barthPrivacyParadoxInvestigating2017; @gerberExplainingPrivacyParadox2018; @kokolakisPrivacyAttitudesPrivacy2017].
@kokolakisPrivacyAttitudesPrivacy2017 come to the conclusion that "the dichotomy between privacy attitude and behaviour should not be considered a paradox anymore." (p. 130)
However, the authors also note that the privacy paradox is a "complex phenomenon that has not been fully explained yet".
@barthPrivacyParadoxInvestigating2017 are more skeptical, and argue that "attempts to theoretically explain and practically solve the problem of the privacy paradox are still scarce and we feel the subject deserves far more research attention" (p. 1052).

<!-- In conclusion, large parts of the current literature suggests that a significant relation between concerns about online privacy and the online sharing of personal information exists and that it is small---which speaks against the privacy paradox. -->

@@

Privacy is defined as the "[...] voluntary and temporary withdrawal of a person from the general society through physical or psychological means [...]" [@westinPrivacyFreedom1967, 7]. 
Hence, privacy captures aspects of both volitional _control_ and social _separateness_.
Several dimensions of privacy have been proposed.
For example, it is possible to distinguish a vertical and a horizontal level.
Whereas the vertical level captures privacy from authorities, institutions, or companies, horizontal privacy addresses privacy from peers, colleagues, or other people [@masurSituationalPrivacySelfdisclosure2018].
When it comes to concerns in general, interestingly they do not seem to be established as a stand-alone theoretical concept in psychology [@colmanDictionaryPsychology2015].
Lexically, concerns are defined as a "marked interest or regard usually arising through a personal tie or relationship" that also reflect an "uneasy state of blended interest, uncertainty, and apprehension" [@merriam-websterDefinitionConcern2018].
Concerns therefore represent both a latent _motivation_ (or increased attention), a negatively valenced _emotion_ (or affective condition), and are mostly implicit.
As a theoretical construct, privacy concerns can hence be categorized as an affective motivational disposition.
As such, there are many similarities with other concepts, including emotions (e.g., fear, anxiety), moods (e.g., dismay, fatigue), attitudes (risk perception, approval), values (e.g., autonomy, freedom), personality traits (e.g., introversion, risk avoidance), and even physiological activation (e.g., attention, arousal).
Taken together, concerns about online privacy represent how much an individual is motivated to focus on his or her control over a voluntary withdrawal from other people or societal institutions on the Internet, accompanied by an uneasy feeling that his or her privacy might be threatened.

The online sharing of personal information, on the other hand, captures how much person-related information people share when they use the Internet, including information about their age, sex, name, address, health, and finances.
Information sharing can be differentiated from communication and self-disclosure.
Communication is broad, because it comprises all verbal and nonverbal information that is emitted [e.g., @watzlawickPragmaticsHumanCommunication2011].
Self-disclosure is narrow, because it focuses on deliberate revelations about the true self to others [e.g., @jourardTransparentSelf1964].
Information sharing is even more specific, because it addresses only person-related information but ignores other types of self-disclosure such as personal fears, values, or plans.

@@

Currently, there is a lack of studies that explicitly analyze how behavior is affected by _concerns_ in general.
Fortunately, however, we know much about the behavioral effects of related concepts such as attitudes or fears, which all can affect behavior profoundly [@fishbeinPredictingChangingBehavior2010; @rogersCognitivePhysiologicalProcesses1983].
Emotions, perhaps the concept most closely related to concerns, have a particularly strong effect on behavior. 
By causing fight, flight, or freeze reactions, they are a primordial trigger of behavior and are considered to be an adaptive mechanism of evolved species [@dolanEmotionCognitionBehavior2002].
<!-- With their direct link to the amygdala, emotions can trigger reactions already subcortically [i.e., without activation of the more recently evolved cortical structures; @dolanEmotionCognitionBehavior2002]. -->

Also empirically, concerns have been shown to affect behavior. 
People more concerned about the environment show more environment-related behaviors [@bambergHowDoesEnvironmental2003].
People more concerned about their appearance consume fewer calories [@hayesConcernAppearanceHealth1987].
People more concerned about their bodies engage in more physical exercise [@reelRelationsBodyConcerns2007].
Taken together, it is reasonable to expect that also concerns about online privacy should reflect in the online sharing of personal information.

At the same time, there are some factors that are likely to diminish the relation.
Most prominently, there is the so-called _attitude behavior gap_ [@fishbeinPredictingChangingBehavior2010], which states that people sometimes act against their own attitudes. 
Not everyone concerned about his or her physical health exercises regularly.
The explanation is simple:
Other factors such as subjective norms and perceived behavioral control also determine behavior [@ajzenIntentionsActionsTheory1985], which automatically reduces the impact of attitudes or concerns. 
Specifically, two of the most influential factors that affect online information sharing are (a) strong subjective norms [@heirmanPredictingAdolescentsDisclosure2013] and (b) expected benefits[@krasnovaOnlineSocialNetworks2010]. 
In other words, users often prioritize social support, special offers, or improved services.
@trepteRiskyBehaviorsHow2014 listed several factors that can additionally attenuate the relation:
lack of strength of concerns,
absence of negative personal experiences, 
or situational constraints due to social desirability.
In conclusion, also in the context of the privacy paradox we should not expect to find a perfect relation between attitudes and behaviors.
Expecting a small to moderate relation is more reasonable.

Finally, there are also some methodological explanations why some studies did not detect statistically significant relations.
Researchers are always confronted with the so-called _Duhem-Quine problem_, which holds that it is impossible to test theories in isolation, because empirical tests always rely on auxiliary assumptions [@dienesUnderstandingPsychologyScience2008]. 
In other words, if a psychological experiment fails, we do not know whether the theory is wrong or the questionnaire subpar. 
This tenet is particularly relevant for the privacy paradox: 
Detecting statistical significance for small effects---and in this case, we should expect to find small effects---is more challenging because it means that large samples are necessary to guarantee sufficient statistical power.[^power]
Precisely, in order to be capable of detecting a correlation between privacy concerns and information sharing in 95% of all cases, which @baruhOnlinePrivacyConcerns2017 estimated to be _r_ = -.13, we need a sample of _N_ = `r pwr.r.test(r = -.13, sig.level = .05, power = .95)$n %>%  round(0)` people. 
The reality, however, looks different: 
In their meta-analysis, @baruhOnlinePrivacyConcerns2017 reported a median sample size of _N_ = 300, which can explain why several studies did not find significant effects.

@@

In conclusion, in line with prior research and the within-person rationales presented above, we expect to find a small significant relation between privacy concerns and information sharing, both on the between-person level and the within-person level.

@@

Although short-term effects are likely, it is still unclear whether long-term effects exist as well.
First, when analyzing potential long-term effects, it is important to choose an interval that is both plausible and relevant.
(It makes a large difference whether the effects of alcohol consumption on driving performance are tested after 1 minute, 1 hour, or 1 day.) 
One factor that determines an interval's optimal length is the stability of the variables [@dormannOptimalTimeLags2015].
Privacy concerns and privacy attitudes are predominantly trait-like constructs with high stabilities, which is why they necessitate longer intervals.
Other studies with comparable research questions have used an interval of 6 months [e.g., @valkenburgEffectsInstantMessaging2009], which we consider to be plausible also in this case.

@@

At the same time, the adverse effect of information sharing on privacy concerns seems plausible as well, with two potential outcomes. 
On the one hand, the long-term relation could be negative: 
If people start to share more information online, they might become increasingly aware that their privacy is at risk, thereby stirring concern.
On the other hand, the long-term relation might also be positive:
When people share more personal information online they might become accustomed to doing so, which potentially reduces concern [for example, due to the mere exposure effect; @zajoncAttitudinalEffectsMere1968].
Finally, there could also be no long-term relation at all: 
People might have already become used to sharing information online, which stifles further cognitive or emotional processing.
This rationale is central to so-called _privacy cynicism_ [e.g., @hoffmannPrivacyCynicismNew2016].

@@

It has been argued that privacy attitudes could bridge the gap between concerns and information sharing [e.g., @dienlinPrivacyParadoxRelic2015]. 
In contrast to privacy concerns, privacy attitudes capture a more explicit, fluctuating cognitive appraisal. 
Although both variables are related to information disclosure, attitudes are likely the better predictor. 
This reasoning follows the rational choice paradigm [@simonBehavioralModelRational1955], which maintains that behavior is always at least partially influenced by convictions, attitudes, and cost-benefit analyses. 
<!-- Likewise, also evolutionary psychology suggests that although emotions guide behavior in the long run, it is more adaptive if behavior instead reflects a cognitive appraisal;  -->
<!-- a tenet arguably reflected best by the phylogenetic development of the neocortex or, if you want, the general dominion of the human race. -->
Also empirically, a study of 1,042 youths from Belgium found that the relation between privacy attitudes and disclosure of personal information was strong (_r_ = .56), whereas the relation between privacy concerns and disclosure was only moderate [_r_ = -.29; @heirmanPredictingAdolescentsDisclosure2013].

@@

Concerning the potential long-term relations of privacy attitudes, we are confronted with the same situation mentioned above.
Because we are not aware of research on long-term relations, several scenarios seem plausible.
Attitudes could either have long-term relations or not, and information sharing could either foster privacy attitudes or diminish them.

@@

This study is part of a large-scale project which investigates the development of privacy and self-disclosure, including severaldother variables. 
Other publications linked to the project can be accessed at https://osf.io/y35as/.
The data come from a longitudinal paper-and-pencil questionnaire study, in which a representative sample of the German population (16 years and older) was surveyed on overall five occasions.
The data can be downloaded from http://dx.doi.org/10.7802/1937.
